---
title: "Untitled"
format: html
---

```{r}
library(dplyr)
library(openxlsx)
library(car)
```

#Création dt_anova
```{r}

data <-  read.xlsx(xlsxFile = "data/climate_data/SuppData.xlsx", fillMergedCells = TRUE, colNames = TRUE)

data <- data %>%
  rename(ENVIRONNEMENT = `Environment.name(c)`) %>%
  mutate(ENVIRONNEMENT = toupper(sub("_.*", "", ENVIRONNEMENT))) %>%
  filter(grepl("N\\+", `Nitrogen(d)`))

partenaire <- read.csv("data/landmarks.csv")

dt_anova <- data %>% left_join(partenaire %>% dplyr::select(ENVIRONNEMENT, PARTENAIRE),by = "ENVIRONNEMENT")

summary(dt_anova)

dt_anova <- dt_anova %>% 
  rename(Year = `Year(a)`) %>%
  rename(Loc = `Location(b)`) %>% 
  rename(SY = `Seed.yield(g)`)

#write.csv(dt_anova, file = "R/dt_anova")
  
View(dt_anova)  

```

Données déséquilibrées --> Plan Type III

Year
Location
Latitude
Longitude

Water capacity
WSC_MAX	Maximal water soil content for a given environment (mm)

Nitrogen
AN	Nitrogen amount in the plant aerial parts at bolting  (g  per m²)
ADB	Plant Aerial Dry Biomass at bolting (t.m²)
NNI	Nitrogen Nutrition Index
Ntotal	mineral nitrogen Input (kg.ha-1)
WSC_I	Water soil content at N-input (% of WSC_MAX) (redondante avec WSC max)
DBI	Number of days without rainfall durig the decade preceeding fertilization
DAI	Number of days without rainfall during the decade post fertilization
RO	number of days with Run Off during the decade after N-input (peu variable)
RF	cumulated Rainfall during the decade after N-input (mm)

- POurquoi on fait pas de la PLS plutôt ? 


#Fichier Dt_anova
```{r}
dt_anova <- read.csv("R/dt_anova", row.names = 1)

dt_anova$Year <- factor(dt_anova$Year)
dt_anova$Loc <- factor(dt_anova$Loc)

summary(dt_anova)
```

# Régression linéaire

```{r}
#Sur les variables qualitatives
contrasts(dt_anova$Year) <- contr.sum(n=nlevels(dt_anova$Year)) 
contrasts(dt_anova$Loc) <- contr.sum(n=nlevels(dt_anova$Loc))

Anova(lm(SY ~ Year+Loc+Latitude+Longitude+WSC_MAX+AN+ADB+NNI+Ntotal+WSC_I+DBI+DAI+RO+RF,dt_anova),type="III") 
#trop de paramètres à estimer
#pas assez d’observations
# Year et Loc trop de facteurs, Pour Year (19)

Anova(lm(SY ~ Latitude+Longitude+WSC_MAX+AN+NNI+Ntotal+DBI+RF,dt_anova),type="III") 
Anova(lm(SY ~ Latitude+Longitude+WSC_MAX+NNI+Ntotal+DBI+RF,dt_anova),type="III")
Anova(lm(SY ~ Latitude+Longitude+WSC_MAX+NNI+DBI+RF,dt_anova),type="III")
Anova(lm(SY ~ Longitude+WSC_MAX+NNI+DBI+RF,dt_anova),type="III")
Anova(lm(SY ~ Longitude+WSC_MAX+NNI+DBI,dt_anova),type="III")
Anova(lm(SY ~ Longitude+NNI+DBI,dt_anova),type="III")
Anova(lm(SY ~ Longitude+DBI,dt_anova),type="III")
#aucun effet significatif 

modele_step <- step(lm(SY ~ Latitude+Longitude+WSC_MAX+AN+NNI+Ntotal+DBI+RF,dt_anova),type="III", direction = "both")
summary(modele_step)
#Mon modèle final après AIC minimal : SY ~ Longitude + WSC_MAX + DBI, Mais aucune variable n’est significative au seuil 0.05
#R² = 0.21, donc seulement ~21% de la variance expliquée.
#Longitude = EO

```




